---
title: "jk_720_week7inclass"
author: "Jason Kearsley"
date: "October 29, 2018"
output: html_document
---
```{r}
read_1 <- "CGCGCAGTAGGGCACATGCCAGGTGTCCGCCACTTGGTGGGCACACAGCCGATGACGAACGGGCTCCTTGACTATAATCTGACCCGTTTGCGTTTGGGTGACCAGGGAGAACTGGTGCTCCTGC"
reads <- c("CGCGCAGTAGGGCACATGCCAGGTGTCCGCCACTTGGTGGGCACACAGCCGATGACGAACGGGCTCCTTGACTATAATCTGACCCGTTTGCGTTTGGGTGACCAGGGAGAACTGGTGCTCCTGC", "AAAAAGCCAACCGAGAAATCCGCCAAGCCTGGCGACAAGAAGCCAGAGCAGAAGAAGACTGCTGCGGCTCCCGCTGCCGGCAAGAAGGAGGCTGCTCCCTCGGCTGCCAAGCCAGCTGCCGCTG", "CAGCACGGACTGGGGCTTCTTGCCGGCGAGGACCTTCTTCTTGGCATCCTTGCTCTTGGCCTTGGCGGCCGCGGTCGTCTTTACGGCCGCGGGCTTCTTGGCAGCAGCACCGGCGGTCGCTGGC")
```

1. Drosophila melanogaster
2. The class is character. length(reads) returns 3, the length of the vector. nchar(reads) returns 124.
3.
```{r}
read_1split <- strsplit(read_1, split = '')
for (i in read_1split) {
  print(i)
}
```
4. It is a list. 
```{r}
read_1splitvec <- unlist(read_1split)
str(read_1splitvec)
```
5. 
```{r}
tabler1s <- table(read_1split)
tabler1s / sum(tabler1s)
```
6.
```{r}
basefreq <- function(x) {
  y <- table(x)
  return(y / sum(y))
}
basefreq(read_1split)
```
This works for both lists and vectors.

2.1 
In the first case x is returned because it was defined outside the loop. In the second case x is merely an internal variable within the function and R is unable to call these.

2.2
```{r}
x <- c()

system.time(for (i in 1:10000) {
  x <- c(x, i^2)
})
```
2.3
```{r}
y <- vector(mode = "integer", length = 10000)

system.time(for (i in 1:10000) {
  y[i] <- i^2
})
```
2.4
Presumably using the apply function with an anonymous function to square the elements of 1:10000. I forget how to do this.
